---
title: "cran-comments"
author: "Kim Hendrickx"
date: "24 juli 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Test environments 

###Check Windows environment  (Microsoft Windows 10 Enterprise)

* platform: x86_64-w64-mingw32/x64 (64-bit) 

* R version 3.4.1 Patched (2017-07-01 r72876)

* R CMD check results
    * 0 errors | 0 warnings | 1 note 
```{r}    
#checking compiled code ... NOTE
#Warning in read_symbols_from_dll(so, rarch) :
#  this requires 'objdump.exe' to be on the PATH
#Warning in read_symbols_from_dll(so, rarch) :
#  this requires 'objdump.exe' to be on the PATH
#Warning in read_symbols_from_dll(so, rarch) :
#  this requires 'objdump.exe' to be on the PATH
#File 'curstatCI/libs/x64/curstatCI.dll':
#  Found no calls to: 'R_registerRoutines', 'R_useDynamicSymbols'
```
 
 I was unable to locate read_symbols_from_dll(so, rach) and could not add objdump.exe to the PATH.
 
###Check Linux environment (Ubuntu 16.04 LTS)

* platform: x86_64-pc-linux-gnu (64-bit) 

* R version 3.4.1 (2017-06-30)

* R CMD check results
    * 0 errors | 0 warnings | 0 notes 

###Check Mac environment  (macOS Sierra version 10.12.5)

* platform: x86_64-apple-darwin15.6.0 (64-bit)
 
* R version 3.4.1  (2017-06-30) 

* R CMD check results 
    * 0 errors | 0 warnings | 0 notes 


## Downstream Dependencies
There are currently no downstream dependencies for this package

